---
title: "Data Manipulation"
author: "Diana Hernandez"
date: "2023-09-21"
output: github_document
---

```{r}
library(tidyverse)
```

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv("data/FAS_pups.csv")

pups_df = janitor::clean_names(pups_df)
```

## 'select'

'select' is used to select columns

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, -litter_number)

select(litters_df, group, litter_id = litter_number)

select(litters_df, group, litter_id = litter_number, everything())

select(litters_df, group, starts_with("pups"))
```

'select' vs 'pull': 'pull' takes something from data frame, 'select' can be used to rename a variable

```{r}
select(litters_df, group)

pull(litters_df, group)
```

'rename'

```{r}
rename(litters_df, litter_id = litter_number)

relocate(litters_df, litter_number)
```

Learning Assessment

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## 'filter'

get rid of rows using 'filter'

= this value into this variable

== check if this is true

!= not equal to

| or 

```{r}
filter(litters_df, group == "Con7")

filter(litters_df, group != "Con7")

filter(litters_df, gd0_weight > 20)

filter(litters_df, gd0_weight <= 20)

filter(litters_df, group == "Con7" | group == "Con8")

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, !(group == "Con7"))

filter(litters_df, group %in% c("Con7", "Con8"), gd0_weight > 20)
```

what about missing rows

drop_na gets rid of missing info anywhere

```{r}
drop_na(litters_df)

drop_na(litters_df, gd0_weight)
```


## 'mutate'

this is used to add or change variables. creating new variables that exist within data frame. can modify and override variables in df.

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, group = str_to_lower(group))

mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## 'arrange'

can reorder data in ascending/descending order

```{r}
arrange(litters_df, gd0_weight)

arrange(litters_df, group, gd0_weight)
```


